---
title: "Data Explorer for Our World in Data COVID dataset"
author: "Nick Lauerman"
date: "7/7/2020"
output: 
  word_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6.5
    fig_height: 6.55
    fig_caption: yes
  pdf_document: 
    toc: yes
    toc_depth: 4
    fig_height: 6.5
    fig_width: 6.5
    fig_caption: yes


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r,echo = TRUE, include=FALSE}
library(tidyverse)
library(lubridate)
```

<!-- # Read Data -->
```{r,echo = FALSE, include=FALSE}
epi_info <- read_csv(file = "./data/Our World Data/full_data.csv")
locations <- read_csv(file = "./data/Our World Data/locations.csv")
```
<!-- ## Country Data -->
```{r,echo = FALSE, include=FALSE}
US <- epi_info %>%
  filter(location == "United States") %>%
  filter(new_cases > 0)
China <- epi_info %>%
  filter(location == "China") %>%
  filter(new_cases > 0) %>%
  filter(new_cases < 5000)
World <- epi_info %>%
  filter(location == "World") %>%
  filter(new_cases > 0)
Germany <- epi_info %>%
  filter(location == "Germany") %>%
  filter(new_cases > 0)
Italy <- epi_info %>%
  filter(location == "Italy") %>%
  filter(new_cases >= 0)
UK <- epi_info %>%
  filter(location == "United Kingdom") %>%
  filter(new_cases > 0)
Spain <- epi_info %>%
  filter(location == "Spain") %>%
  filter(new_cases > 0)
Canada <- epi_info %>%
  filter(location == "Canada") %>%
  filter(new_cases > 0) %>%
  filter(new_cases < 2500)
Mexico <- epi_info %>%
  filter(location == "Mexico") %>%
  filter(new_cases > 0)
Ireland <- epi_info %>%
  filter(location == "Ireland") %>%
  filter(new_cases > 0)
CostaRica <- epi_info %>%
  filter(location == "Costa Rica") %>%
  filter(new_cases > 0)
India <- epi_info %>%
  filter(location == "India") %>%
  filter(new_cases > 0)


```

# Plot

## World
```{r, fig.cap= "World wide new cases", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = World) +
  geom_point() 
```
```{r, fig.cap= "World Wide New Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = World) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "World wide total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = World) +
  geom_point() 
```
```{r, fig.cap= "World Wide total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = World) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="World wide new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = World) +
  geom_point()
```
```{r, fig.cap="World wide new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = World) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
``

```{r, fig.cap= "World wide total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = World) +
  geom_point() 
```
```{r, fig.cap= "World Wide total deaths with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = World) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```


## Countries of Interest
### US
```{r, fig.cap= "US", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = US) +
  geom_point() 
```
```{r, fig.cap= "US with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = US) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "US total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = US) +
  geom_point() 
```
```{r, fig.cap= "US total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = US) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="US new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = US) +
  geom_point()
```
```{r, fig.cap="US new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = US) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "US total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = US) +
  geom_point() 
```
```{r, fig.cap= "US total deaths with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = US) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```

### Germany
```{r, fig.cap= "Germany", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Germany) +
  geom_point() 
```
```{r, fig.cap= "Germany with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Germany) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Germany total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Germany) +
  geom_point() 
```
```{r, fig.cap= "Germany total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Germany) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="Germany new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Germany) +
  geom_point()
```
```{r, fig.cap="Germany new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Germany) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Germany total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = Germany) +
  geom_point() 
```

### India

```{r, fig.cap= "India", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = India) +
  geom_point() 
```
```{r, fig.cap= "India with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = India) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "India total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = India) +
  geom_point() 
```
```{r, fig.cap= "India total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = India) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="India new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = India) +
  geom_point()
```
```{r, fig.cap="India new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = India) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "India total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = India) +
  geom_point() 
```

## Other Countries
### China
```{r, fig.cap= "China", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = China) +
  geom_point() 
```
```{r, fig.cap= "China with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = China) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "China total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = China) +
  geom_point() 
```
```{r, fig.cap= "China total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = China) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="China new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = China) +
  geom_point()
```
```{r, fig.cap="China new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = China) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "China total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = China) +
  geom_point() 
```

### Italy
```{r, fig.cap= "Italy", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Italy) +
  geom_point() 
```
```{r, fig.cap= "Italy with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Italy) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Italy total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Italy) +
  geom_point() 
```
```{r, fig.cap= "Italy total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Italy) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="Italy new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Italy) +
  geom_point()
```
```{r, fig.cap="Italy new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Italy) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Italy total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = Italy) +
  geom_point() 
```

### Spain
```{r, fig.cap= "Spain", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Spain) +
  geom_point() 
```
```{r, fig.cap= "Spain with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Spain) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Spain total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Spain) +
  geom_point() 
```
```{r, fig.cap= "Spain total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Spain) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="Spain new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Spain) +
  geom_point()
```
```{r, fig.cap="Spain new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Spain) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Spain total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = Spain) +
  geom_point() 
```

### Mexico
```{r, fig.cap= "Mexico", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Mexico) +
  geom_point() 
```
```{r, fig.cap= "Mexico with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Mexico) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Mexico total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Mexico) +
  geom_point() 
```
```{r, fig.cap= "Mexico total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Mexico) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="Mexico new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Mexico) +
  geom_point()
```
```{r, fig.cap="Mexico new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Mexico) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Mexico total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = Mexico) +
  geom_point() 
```

### Canada
```{r, fig.cap= "Canada", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Canada) +
  geom_point() 
```
```{r, fig.cap= "Canada with trend line", echo=FALSE}
ggplot(aes(x = date, y=new_cases), data = Canada) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Canada total cases", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Canada) +
  geom_point() 
```
```{r, fig.cap= "Canada total Cases with trend line", echo=FALSE}
ggplot(aes(x = date, y=total_cases), data = Canada) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap="Canada new deaths", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Canada) +
  geom_point()
```
```{r, fig.cap="Canada new deaths with trending", echo=FALSE}
ggplot(aes(x = date, y=new_deaths), data = Canada) +
  geom_point() +
  geom_smooth(method = "loess",
              formula = y~x)
```
```{r, fig.cap= "Canada total deaths", echo=FALSE}
ggplot(aes(x = date, y=total_deaths), data = Canada) +
  geom_point() 
```
